---
title: "RQ1 summary code"
author: "Erica Zaja"
date: '2022-04-14'
output: pdf_document
---

# # RQ1: How is shrub biomass distributed in the focal study area?

### Part 1: BUFFERED RANDOM SAMPLING 


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message= FALSE}
## LOADING LIBRARIES -----
library(sp)
library(raster)
library(ggplot2)
library(viridis)
library(rasterVis)
library(sf)
library(tidyverse)
```


```{r, warning = FALSE}
## LOADING DATA ----

# SHRUB DATA: from Berner et al 2018
# Loading raster of shrub biomass (g/m2) on Alaskan North Slope  
shrub_agb_p50 <- raster("../../datasets/berner_data/shrub_agb_p50.tif") 
# Using the best-estimates: the 50th percentile of the 1,000 permutations

# PCH CORE RANGE DATA: from Porcupine Caribou Management Board (2016)
# Loading polygon of PCH range 
PCH_core_range <- st_read("../../datasets/PCH_Core_Range_2016/PCH_Core_Range_2016.shp") # loading data

```

```{r}
## EXPLORING DATA ------

# class: what type of data 
class(shrub_agb_p50) # RasterLayer
class(PCH_core_range) # sf dataframe

# resolution of shrub map 
res(shrub_agb_p50) # resolution of map: [1] 30m x 30m

```

```{r}
## CROPPING -----

# Cropping shrub raster to the PCH range 
#r2 <- crop(shrub_agb_p50, extent(PCH_core_range))
#r3 <- mask(r2, PCH_core_range)
#plot(r3) # plot cropped raster
#plot(PCH_core_range, add = TRUE, lwd = 2)

# transforming CRS of cropped map from proj = aea (alaska albers) to proj = latlong 
# r3_latlong <- projectRaster(r3, crs="+init=EPSG:4326", xy = TRUE)

```

```{r}
## AGGREGATION ----
r3_latlong <- raster("../../datasets/berner_data/r3_latlong.tif")

# aggregate shrub data to coarser resolution before extraction using aggregate()
# factor chosen dividing climate cell resolution 0.008333333 x 0.008333333 by 
# the resolution of the cropped shrub map (latlong)
# r3_latlong_agg <- aggregate(r3_latlong, fact=c(11.47842,30.8642), fun = mean, 
# expand = TRUE) 

# Loading aggregated raster
r3_latlong_agg <- raster("../../datasets/berner_data/r3_latlong_agg.tif")

```

```{r, error = FALSE, warning = FALSE}
# RANDOM SAMPLE WHOLE MAP ----

# Measuring area of raster
# get sizes of all cells in raster [km2]
cell_size <- raster::area(r3_latlong_agg, na.rm=TRUE, weights=FALSE)
# delete NAs from vector of all raster cells
# NAs lie outside of the rastered region, can thus be omitted
cell_size <- cell_size[!is.na(cell_size)] # 0.2815663
#compute area [km2] of all cells in geo_raster
raster_area <-length(cell_size)*median(cell_size)
# print area of shrub map according to raster object
print(paste("Area of PCH Alaskan range (raster)", round(raster_area, digits = 1),
            "km2"))


# Buffered random sampling 

# c. Extracting 1 every 3 pixels (9583/3 = 3195)
# r3_rsample_1 <- as.data.frame(sampleRandom(r3_latlong_agg, 3195, buffer = 1414.2, 
#na.rm=TRUE, ext=NULL, cells=TRUE, rowcol=FALSE, xy = TRUE)) 

# Checking buffer works
# calculating Haversine distance between points (x and y coordinates)
# r3_rsample_01 <- r3_rsample_1  %>% 
 # mutate(r3_rsample_1, Distance = distHaversine(cbind(x, y),
                                                #cbind(lag(x), lag(y))))

# If distance between points > buffer distance, buffer works
#r3_rsample_01 <- r3_rsample_01 %>% 
 # mutate(buff = case_when(Distance >= 1414.2 ~ "T", Distance < 1414.2 ~ "F"))

# r3_rsample_01 <- r3_rsample_01 %>%  filter(buff %in% c("T")) 
# only keeping observations where buff worked

# Clean random sample dataset saved as r3_rsample_002
r3_rsample_002 <- read_csv("../../datasets/berner_data/r3_rsample_002.csv")

```



```{r, fig.show="hide" }
## DATA VISUALISATION -----

# Plotting cropped map with personalised color palette
# Colour palette credit: David Nichols (WONG colour blind friendly palette)

# making raster into dataframe
r3_latlong_agg_df <- as.data.frame(r3_latlong_agg, xy=TRUE, 
                                   na.rm=TRUE)

# setting biomass level thresholds using quantiles 
r3_latlong_agg_df <- r3_latlong_agg_df %>%
  mutate(biomass_level = case_when (r3_latlong_agg < 170.630482 ~ 'Low', 
                                    r3_latlong_agg> 170.630482    
                                    & r3_latlong_agg < 347.062210 ~ 'Medium', 
                                    r3_latlong_agg > 347.062210 ~ 'High'))

# ordering factor levels
r3_latlong_agg_df$biomass_level <- factor(r3_latlong_agg_df$biomass_level,
                                          levels=c("Low", "Medium", "High"),
                                         labels = c("Low", "Medium", "High"),
                                         ordered = T)

# plotting raster with personalised colours
(raster_my_palette_new <- ggplot(r3_latlong_agg_df) + 
  geom_tile(aes(x=x,y=y,fill=biomass_level)) + 
  scale_fill_manual(name = "Biomass level", values=c("#F0E442", "#E69F00", 
                                                      "#009E73")) +
  coord_quickmap()+
 # theme_shrub() +  
  xlab("\nLongitude") +
  ylab("Latitude\n") +
  xlim(-146.5, -141)+
  ylim(69.2,70.2)+ 
  theme(plot.title = element_text(hjust = 0.5),      
                       text = element_text(size=25),	
        axis.title.x = element_text(size=25),
        axis.title.y = element_text(size=25),
        axis.text.x = element_text(size=25, hjust = 1),
                       axis.text.y = element_text(size=25, hjust = 1),
                       legend.text = element_text(size=20),
                       legend.title = element_text(size=30), 
                       legend.position ="bottom"))
```


### Part 2: MODELLING
```{r pressure, echo=FALSE}
## DATA MANIPULATION ----

# Exploring data
hist(r3_rsample_002$biomass) # distribution ```

```


```{r, echo=FALSE}
##  MODELLING ----
# Standardising lat and long (explanatory variables)
r3_rsample_002$latitude <- scale(r3_rsample_002$latitude, center = TRUE, 
                                 scale = TRUE)
r3_rsample_002$longitude <- scale(r3_rsample_002$longitude, center = TRUE, 
                                  scale = TRUE)

# MODEL 1. biomass vs lat ----
model_1 <- lm(biomass~latitude, data = r3_rsample_002)
summary(model_1)

# Checking model 1 assumptions 
plot(model_1)
qqnorm(resid(model_1))
qqline(resid(model_1))  # points fall nicely onto the line - good!


```


```{r, echo=FALSE, fig.show="hide"}
# Extracting model predictions 
predictions_1 <- as.data.frame(predict(model_1, newdata = r3_rsample_002, 
                                       interval = "confidence")) 
model_1_lat <- cbind(r3_rsample_002, predictions_1)

# Plot the predictions 
(predictions_biomass_vs_lat <- (ggplot(model_1_lat, aes(latitude, fit)) + 
                      geom_point(data = model_1_lat, aes(x = latitude, y = biomass), 
                                 colour =  "#006146", size = 0.5) +
                      stat_smooth(method=lm, colour = "black", size = 2)+
                      geom_line(aes(y=lwr),  color = "#F96E00", 
                                linetype = "dashed", size = 0.5)+
                      geom_line(aes(y=upr), color = "#F96E00", 
                                linetype = "dashed", size = 0.5)+
                        annotate(geom = "text", x = 1, y = 800, label="slope = -49.079*** ",
                                 size = 10) +
                      xlab("\nScaled latitude") +
                      ylab(bquote("Shrub biomass "*(g~m^-2)*""))+ 
                      # theme_shrub()+
                      theme(axis.title.x =element_text(size=25),
                              axis.title.y =element_text(size=25),
                              axis.text.x = element_text(size=25, hjust = 1),
                              axis.text.y = element_text(size=25, hjust = 1) )) )

